---
title: 'Radiocarbon Course: Isotopic Partitioning Models'
author: "Marguerite Mauritz"
date: "21 August 2021"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---

#Broad Questions
1. What is the relative contribution from plant (above-ground and below-ground) and soil (young and old) sources to total ecosystem respiration, using ^14^C to separate sources?  
2. How does the addition of ^13^C isotopes affect the partitioning model relative contirbutions to respiration? (plant above-and below-ground, soil young and old) 


#The Data
Data from this exercise is based on [Pegoraro et al 2020](https://doi.org/https://doi.org/10.1111/gcb.15481).  

Download and save the following data files in the same directory (folder) as the Rmd code. R will read files from the same path and you will not have to change the working directory: 

```{r, create a table for data files to import, echo=FALSE}
# won't display in-line in Rmd but will knit to html

# create data to display file info
table.data <- data.frame(File.Name=c("Model_14C_output.csv", "Model_14C_13C_output.csv"),
                         File.Description=c("Plot-level relative proportions of AG, BG, OS, YS using 14C partitioning", "Plot-level relative proportions of AG, BG, OS, YS using 14C and 13C partitioning"))

# print table with nicer header names
knitr::kable(table.data, col.names=gsub("[.]", " ", names(table.data)))

```

# Activity Preparation: Load libraries
If these libraries are not installed on your computer, install them using `install.packages(dplyr)`,`install.packages(ggplot2)`, etc. or manually install them from the 'Packages' tab in RStudio.  
```{r, load or install packages, message=FALSE}
library(dplyr) # organise and examine data
library(ggplot2) # make plots
library(gridExtra) # for combining final plots at the end
```


# Part 1: Relative Source Contributions in the Single ^14^C Isotope Model

## Import data
For part 1, import the data from the single isotope model.  
Use `model.single <- read.csv("FILE NAME.csv")` to load the correct dataset.  
```{r, import the data model.single, echo=TRUE}
# add name of 
model.single <- read.csv("Model_14C_output.csv")
```

## Examine data structure

The model outputs estimated variable values in seperate rows, each row has a name and index according to the parameter definitions in the model. The data is organised by variable name, in alphabetical order.  
To use the output, we need to extract relevant parameters based on their row position. 

```{r, look at data structure for model.single}
# use glimpse() to look at the data structure
glimpse(model.single)
```

```{r, look at the variable names for model.single}
# look at the variable names and orders by calling only the variable column
model.single$variable
```

##Orient to the data

#####a. Refer back to the model definition, what does each variable value and row represent?  
A:  
B:  
etc.  

#####b. What do the columns mean, UL, LL represent?  

## Extract the relative AG, BG, OS, YS proporation estimates for each plot
Use the row indices in model.single to extract only the data rows that estimate the relative flux contributions. Combine them into a single data frame using `rbind()`.  
```{r, create dataframe of relative proportions in model.single, echo=TRUE}
proportions.single <- rbind(model.single[22:63,],
                            model.single[65:106,],
                            model.single[108:149,],
                            model.single[151:192,])
```

Now create a column called 'source' to group all observations from a single source under the same name and begin plotting the data. The function [grepl()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/grep) searches for pattern matches in a string of characters.  
```{r, create a source column in proportions.single to group sources}
proportions.single <- proportions.single %>%
  mutate (source = case_when(grepl("PAG", variable) ~ "Aboveground",
                             grepl("PBG", variable) ~ "Belowground",
                             grepl("POS", variable) ~ "OldSoil",
                             grepl("PYS", variable) ~ "YoungSoil"))

```

## Graph the plot-level estimates 
To visualise the individual obserations, graph the means of the plot-level relative source contributions for each plot, grouped by source:

```{r}
ggplot(proportions.single, aes(source, mean))+
  geom_point()
```

## Graph the mean estimates
The model also calculated mean proportions of each source denoted by _mean. Extract the four mean estimates from **model.single**, create a 'source' column as in proportions.single, and graph them together with the upper and lower credible intervals. 

```{r}
mean.proportions.single <- rbind(model.single[64,],
                            model.single[107,],
                            model.single[150,],
                            model.single[193,])

mean.proportions.single <- mean.proportions.single %>%
  mutate (source = case_when(grepl("PAG", variable) ~ "Aboveground",
                             grepl("PBG", variable) ~ "Belowground",
                             grepl("POS", variable) ~ "OldSoil",
                             grepl("PYS", variable) ~ "YoungSoil"))

ggplot(mean.proportions.single, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))

```

## Create a nice looking figure of the mean proportions and credible intervals
Use `+ylim()` in the ggplot to fix the y-axis limits from 0 to 1 and `+labs()` to add a title, x-axis, and y-axis labels. You can also use `'theme_bw()` or other default theme to make the overall plot appearance a little nicer. 

```{r, plot mean source estimates from model.single, eval=TRUE, echo=TRUE}
fig1 <- ggplot(mean.proportions.single, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))+
  ylim(0,1)+
  labs(x="Respiration Source",y="Relative mean proportion",title="Single 14C Isotope source contribution estimates")+
  theme_bw()

fig1
```

# Part 2: Relative Source Contributions in the Dual ^14^C and ^13^C Isotope Model

## Import data
For part 2, import the data from the dual isotope model.  
Use `model.single <- read.csv("FILE NAME.csv")` to load the correct dataset.  
```{r, import the data for model.dual, echo=TRUE}
# add name of 
model.dual <- read.csv("Model_14C_13C_output.csv")
```

## Examine data structure

The model outputs estimated variable values in seperate rows, each row has a name and index according to the parameter definitions in the model. The data is organised by variable name, in alphabetical order.  
To use the output, we need to extract relevant parameters based on their row position. 

```{r, look at data structure for model.dual}
# use glimpse() to look at the data structure
glimpse(model.dual)
```

```{r, look at the variable names for model.dual}
# look at the variable names and orders by calling only the variable column
model.dual$variable
```
## Extract the relative AG, BG, OS, YS proporation estimates for each plot from the dual model
Use the row indices in model.single to extract only the relative flux contributions. Combine them into a single data frame using `rbind()`.  
Create an object **proportions.dual <-** as above.

```{r, create dataframe of relative proportions in model.dual}
proportions.dual <- rbind(model.dual[21:62,],
                            model.dual[64:105,],
                            model.dual[107:148,],
                            model.dual[150:191,])
```

Now create a column called 'source' to group all observations from a single source under the same name and begin plotting the data. The function [grepl()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/grep) searches for pattern matches in a string of characters.  
```{r, create a source column in proportions.dual to group sources}
proportions.dual <- proportions.dual %>%
  mutate (source = case_when(grepl("PAG", variable) ~ "Aboveground",
                             grepl("PBG", variable) ~ "Belowground",
                             grepl("POS", variable) ~ "OldSoil",
                             grepl("PYS", variable) ~ "YoungSoil"))

```

## Graph the plot-level estimates 
To visualise the individual obserations, graph the means of the plot-level relative source contributions for each plot, grouped by source:

```{r}
ggplot(proportions.dual, aes(source, mean))+
  geom_point()
```

## Graph the mean estimates
The model also calculated mean proportions of each source denoted by _mean. Extract the four mean estimates from **model.single**, create a 'source' column as in proportions.single, and graph them together with the upper and lower credible intervals. 

```{r}
mean.proportions.dual <- rbind(model.dual[63,],
                            model.dual[106,],
                            model.dual[149,],
                            model.dual[192,])

mean.proportions.dual <- mean.proportions.dual %>%
  mutate (source = case_when(grepl("PAG", variable) ~ "Aboveground",
                             grepl("PBG", variable) ~ "Belowground",
                             grepl("POS", variable) ~ "OldSoil",
                             grepl("PYS", variable) ~ "YoungSoil"))

ggplot(mean.proportions.dual, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))

```

## Create a nice looking figure of the mean proportions and credible intervals
Use `+ylim()` in the ggplot to fix the y-axis limits from 0 to 1 and `+labs()` to add a title, x-axis, and y-axis labels. You can also use `'theme_bw()` or other default theme to make the overall plot appearance a little nicer. 

```{r, plot mean source estimates from model.dual, eval=FALSE, echo=FALSE}
fig2 <- ggplot(mean.proportions.dual, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))+
  ylim(0,1)+
  labs(x="Respiration Source",y="Relative mean proportion",title="Dual 14C & 13C Isotope source contribution estimates")+
  theme_bw()

fig2
```

# Part 3: Compare the outputs of the single and double isotope model relative proportion estimates

##Write a brief summary comparing the outputs from the single and dual isotope partitioning models, using the four prompts below. 

###a. Print the nicely formated plots showing the mean source contributions from the two different models

```{r, print single and dual model plots, echo=FALSE}

fig1 <- ggplot(mean.proportions.single, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))+
  ylim(0,1)+
  labs(x="Respiration Source",y="Relative mean proportion",title="Single 14C Isotope \n source contribution estimates")+
  theme_bw()

fig2 <- ggplot(mean.proportions.dual, aes(source, mean))+
  geom_point()+
  geom_pointrange(aes(ymin=LL,ymax=UL))+
  ylim(0,1)+
  labs(x="Respiration Source",y="Relative mean proportion",title="Dual 14C & 13C Isotope \n source contribution estimates")+
  theme_bw()

grid.arrange(fig1,fig2,ncol=2)
```


###b. Report the the group means and credible intervals, and the min/max range of the plot-level estimates for each source, from the two models.

###c. What does the addition of the ^13^C isotope change in the relative source estimates? Why?   
*It might help to know that in the Bayesian models, two groups are considered significantly different if the mean of one group is not contained by the credible interval of another*

###d. What is the benefit of having additional isotopes in the partitioining model? 
